---
title: "Disclosure raises prices where search costs are high"
subtitle: "Seminar, Department of Economics"
author: "Your Name"     # the exact form you want on the title slide; see style/house.md
institute: "Your University"
date: last-modified
format: starter-revealjs
---

## Notation {visibility="hidden"}

\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\mathrm{Var}}
\DeclareMathOperator*{\argmin}{arg\,min}

```{r setup}
#| include: false
library(ggplot2)
# load the analysis objects the deck needs
```

## Mandated disclosure was supposed to lower prices

Forty states passed it in six years. Prices rose in half of them.

::: {.aside-note}
Source: Author (2024), Table 2.
:::

::: {.notes}
The policy rationale, the two states everyone cites, and why the puzzle went
unnoticed. Say this; do not put it on the slide.
:::

## {.center}

::: {.r-fit-text}
The sign flips
:::

## Where search costs are high, disclosure raises prices instead

::: {.result}
[Estimate]{.label}
0.29 log points (s.e. 0.04) in the top search-cost quartile, 0.06 (0.05)
elsewhere.
:::

::: {.takeaway}
The sign of the policy effect flips with market structure.
:::

## Setting {.section-break}

## The reform hit half the markets in 2018 and none of the others

- 1,240 markets, 2014 to 2022
- Treatment assigned by a statutory population cutoff
- Prices from weekly scanner data

## Identification compares high- and low-search-cost markets before and after

$$
\log p_{it} = \alpha_i + \gamma_t + \beta\,(\text{Post}_t \times \text{High}_i) + \varepsilon_{it}
$$

::: {.assumption}
[Assumption 1]{.label}
Absent the reform, prices in high- and low-search-cost markets would have moved
in parallel.
:::

## Prices diverge only after the reform

```{r}
#| fig-width: 8
#| fig-height: 4.2
# event-study plot; annotate the focal coefficient in the figure itself
```

## The estimate survives every control set

:::: {.columns}
::: {.column width="58%"}
```{r}
#| fig-width: 5.4
#| fig-height: 4.0
# coefficient plot across specifications
```
:::
::: {.column width="42%"}
Point estimate moves by less than one standard error across specifications.

::: {.result}
[Estimate]{.label}
0.29 (0.04)
:::

::: {.aside-note}
Low-search-cost markets move by 0.06, not distinguishable from zero.
:::

::: {.with-previous}
[Sample Windows]{.jump target="ap-window"}
:::
:::
::::

## Search costs explain the reversal; concentration does not

::: {.fragment .fade-in-then-semi-out}
Effect scales with baseline search cost.
:::

::: {.fragment .fade-in-then-semi-out}
No gradient in concentration.
:::

::: {.fragment}
Both hold within state and within year.
:::

## Disclosure helps only where consumers can act on it

::: {.result}
[Estimate]{.label}
0.29 log points (s.e. 0.04) where search costs are high.
:::

The claim, restated in the words you want quoted.

## {.thanks-slide .center .no-stage}

::: {.thanks-title}
Thank You!
:::

::: {.thanks-invite}
Comments and questions welcome:
:::

::: {.thanks-contact}
[you@example.edu](mailto:you@example.edu)

[your-site.example](https://your-site.example)
:::

::: {.qr-slot}
Paper
:::

## Appendix {.appendix-break background-color="var(--appendix-ground)"}

## Robustness to the sample window {.appendix #ap-window}

| Specification | Estimate | s.e. |
|---|---:|---:|
| Baseline | 0.291 | 0.041 |
| Drop 2020 | 0.284 | 0.044 |
| Market-specific trends | 0.276 | 0.049 |

: Sensitivity of the main estimate {.spec}

::: {.aside-note}
Nothing moves the estimate by half a standard error.
:::

::: {.with-previous}
[]{.jump-back}
:::

## Why not an instrument for search costs {.appendix #ap-iv}

Pre-built answer to the question you expect.

::: {.with-previous}
[]{.jump-back}
:::

<!-- Only on a deck that cites: add `bibliography: talk-refs.bib` to the front
matter and the references block from references/citations.md. -->

<!-- On this file as-is, the placeholder chunks draw nothing, so the fit gate
warns three slides nearly empty and stage-check fails the event-study slide;
both resolve once real exhibits fill the chunks. The quote slide reports
HEADING NOT VISIBLE either way; see the traps list in SKILL.md. -->
